---
title: "GSS challange Less Data"
author: "Judy Insunza"
date: "10/16/2025"
output: html_document
---



```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??





# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

This problem was a categorical problem. We found out in the categorical of race, there was a factor with 3 levels. Those were called "white", "black" and "others." Cappun is a categorical variable with 2 levels which are in favor of capital punishment or oppose capital punishment. These were called "favor" and "oppose."

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->





```{r}
dd3 <- mydata %>% group_by(cappun,race) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The white people are for it with 2 to 1. The black people are more aginst it, and other seem more for it.
## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.



### Numerical Descriptive Results
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~race+ cappun, data=mydata)


```

The majority of white people favor capital punishment, with 67.75% supporting it. Black people only had 47.24% and other had a significant value of 61.25%. The majority of black people opposed capital punishment, with a value of 52.76%. White people only opposed it by 38.75%.

Of the people who favor capital punishment, almost 80% of them are white people, black people made up 11.72% of the people who favored capital punishment 8.47% were other people. In the opposing side was 65.38%, which were white. Lastly, there were was 10.51% who were others. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
fisherXY(mydata$cappun,mydata$race,printAll = TRUE)
```









```{r}
fisherXY (mydata$race,mydata$cappun)
```

Statistically, every group is different than every group. With a low-value each of the groups.

When I observed the inferential results, it gave me the P value from comparison between White and black. White and other, and other. For White black the P-value was so low 1.741e05 which showed that they are statistically different, but for the white vs others and black vs others the P-values were 0.2565 and 0.06299 which shows us that they are not statistically different. However, in the mastor branch, when doing the fisher test it showed us that all the values were statistically different

You can mostly see that see that black and white are the ones with most difference, according to the p-values. it has a very low number of around 0.000017. More white people are for the capital punishments, while less black people are for the capital punishment. However, between white and other, the p-value is only 0.2565. This means that there really isn't a difference between the amount of people favor capital punishment for white and others. They both prefer capital punishment. Lastly, the p-value between black and others is 0.06299. This shows that they are also close. It is less than p-value of white and others, but it is significantly not different. Black people are more aginst capital punishment then others.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

While this process was similar to the one done in the master branch, there is a significantly different amount of observations. This is due to the fact that we only concluded a statistically difference between the black & white. This only gave us 821 observations. This led to there not being much difference in p-values,expect for the black and white. There is less to go off of, when you have less observations. Which was backed up by what I observed during fisher's test in the number 2 branch.  You can mostly see that see that black and white are the ones with most difference, according to the p-values. it has a very low number of around 0.000017. More white people are for the capital punishments, while less black people are for the capital punishment. However, between white and other, the p-value is only 0.2565. This means that there really isn't a difference between the amount of people favor capital punishment for white and others. They both prefer capital punishment. Lastly, the p-value between black and others is 0.06299.This shows that they are very close. The graphical and numerical descriptive results showed the same thing. They both showed percentage & visual values of the black and white different because black people are against capital punishment, and white others being almost the same because they both support capital punishment more.
